---
title: "SDS 315 Homework 3"
author: "Brian Hu (bth997)"
date: "2/11/2025"


output: 
  html_document: 
    toc: true
    toc_float: true
    toc_collapsed: true
    toc_depth: 3
    theme: lumen
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(tidyverse)
library(ggplot2)
library(kableExtra)
library(mosaic)
```

Link to github respository with Rmd file: <https://github.com/brianbevo/SDS_315/tree/main/HW3>

## Question 1

```{r, echo=FALSE, message=FALSE}

setwd('/Users/brianhu/Desktop/SDS_315')
gasprice <- read_csv("gasprices.csv", show_col_types = FALSE)
```

### Theory A - Gas Prices charge more if they lack direct competition in sight

```{r, echo=FALSE, message=FALSE}

boot_competition <- do(10000)*diffmean(Price ~ Competitors, data = resample(gasprice))

print(confint(boot_competition, level = 0.95))

mean(Price ~ Competitors, data= gasprice)

```

We are trying to see if a lack of competitive gas stations in sight changes or charges the gas price higher.The evidence behind this claim does not support the argument as the the 95% confidence interval has 0 which means this results are not statistically significant. Also the difference in price is only around 2 cents, emphasizing that this theory is not supported by our data.

### Theory B - The richer the area, the higher the gas prices

```{r, echo=FALSE, message=FALSE}

income <- do(10000)*cor(Price ~ Income, data = resample(gasprice))

print(confint(income, level = 0.95))

```

We are trying to see if when the income of an area is higher it influences and raises gas prices in the area. The correlation coefficient estimate shows a value of .396 meaning that the evidence supports a statistically significant difference and that there is might be strong correllation, so it can can be said that the theory has data to support it.

### Theory C - Gas stations at stoplights charge more

```{r, echo=FALSE, message=FALSE}

stop <- do(10000)*diffmean(Price ~ Stoplight, data = resample(gasprice))
                                  
print(confint(stop, level = 0.95))

mean(Price ~ Stoplight, data= gasprice)
```

We are trying to see if gas stations at stop lights charge the gas price higher than those not at stop lights.The evidence behind this claim does not support the argument as the the 95% confidence interval has 0 which means this results are not statistically significant. Also the difference in price is only around 0 cents, emphasizing that this theory is not supported by our data.

### Theory D - Gas stations with direct highway access charge more

```{r, echo=FALSE, message=FALSE}
boot_highway <- do(10000)*diffmean(Price ~ Highway, data = resample(gasprice))
                                  
print(confint(boot_highway, level = 0.95))

mean(Price ~ Highway, data= gasprice)
```

We are trying to see if gas stations at highways with direct access charge the gas price higher than those not at direct highway access lights.The evidence behind this claim does support the argument as the the 95% confidence interval does not include the value 0 which means these results are statistically significant. Also the difference in price is around 4.6 cents, emphasizing that this theory is supported by our data.

### Theory E - Shell charges more than all other non-Shell brands.

```{r, echo=FALSE, message=FALSE}

shell_boot = do(10000)*diffmean(Price ~ Brand=="Shell", data=resample(gasprice))
                                   
print(confint(shell_boot, level = 0.95))

mean(Price ~ Brand=="Shell", data= gasprice)
```

We are trying to see if gas stations under the brand Shell charge higher than other brands in the same market .The evidence behind this claim does not support the argument as the the 95% confidence interval has 0 which means this results are not statistically significant. Also the difference in price is only around 2.7 cents, emphasizing that this theory is not supported by our data given.

## Question 2

### Part A

```{r, echo=FALSE, message=FALSE}

cars <- read_csv("sclass.csv", show_col_types = FALSE)
cars_a <- filter(cars, year == 2011, trim == "63 AMG")

cars_a_boot  <- do(10000)*mean(~mileage, data = resample(cars_a))
confint(cars_a_boot, level = 0.95)

```

After the confidence test, we are 95% confident that true mileage mile amount for the 201 S-Class 63 AMGS can be found between the interval from 2625.26 to 3823.72 which also has an estimate value of 28997.34.

### Part B

```{r, echo=FALSE, message=FALSE}

cars_b <- filter(cars, year == 2014, trim == "550")

cars_b_boot = do(10000)*prop(~ color=="Black", data=resample(cars_b ))
confint(cars_a_boot, level = 0.95)

```

After the confidence test, we are 95% confident that there were around 41% to 46% of 2014 S-Class 550s that have a black exterior.

## Question 3

### Part A

```{r, echo=FALSE, message=FALSE}
  
nbc <- read_csv("nbcpilotsurvey.csv", show_col_types = FALSE)

clean_nbc_a <- filter(nbc, Show == "Living with Ed" | Show == "My Name is Earl")

nbc_a_boot <- do(10000)*diffmean(Q1_Happy ~ Show, data = resample(clean_nbc_a))

confint(nbc_a_boot, level = 0.95)

```

The question tasked today is determining whether the shows "Living with Ed" and "My name is Earl" have different audience happiness approval ratings and what that might mean. My first step was clean the data by filtering out between which show the response was for; then taking the cleaned data, I applied a bootstrap of 10,000 trials whilst deploying a confidence interval of 95% to observe if there is a significant difference within the opposing happiness ratings. The bootstrap showed that after 10,000 trials the confidence interval ranged from -0.400 to 0.104 which included the value 0. Since our interval has the value 0 we can say there is not a statistically significant difference that one of the shows is repeatedly causing a higher average happy response from its audience.

### Part B

```{r, echo=FALSE, message=FALSE}
  
clean_nbc_b <- filter(nbc, Show == "The Biggest Loser" | Show == "The Apprentice: Los Angeles")

nbc_b_boot <- do(10000)*diffmean(Q1_Annoyed ~ Show, data = resample(clean_nbc_b))

confint(nbc_b_boot, level = 0.95)

```

The question tasked today is determining whether the shows "The Biggest Loser" and "The Apprentice: Los Angeles" have different audience annoyance disapproval ratings and what that might mean. My first step was clean the data by filtering out between which show the response was for; then taking the cleaned data, I applied a bootstrap of 10,000 trials whilst deploying a confidence interval of 95% to observe if there is a significant difference within the opposing happiness ratings. The bootstrap showed that after 10,000 trials the confidence interval ranged from -0.520 to -0.022 which does not include the value 0. Since our interval does not have the value 0 we can say there is a statistically significant difference that one of the shows is repeatedly causing a higher average annoyance response from its audience. When taking the difference in means of ratings between the show we can see that there on a scale from 1 - 5, an average annoyance rating for "The Apprentice: Los Angeles" was around 0.271 higher than an annoyance rating from "The Biggest Loser".

### Part C

```{r, echo=FALSE, message=FALSE}

clean_nbc_c <- filter(nbc, Show == "Dancing with the Stars")
nbc_c_boot = do(10000)*prop( ~ Q2_Confusing >= 4, data=resample(clean_nbc_c))

confint(nbc_c_boot, level = 0.95)

```

The question tasked today is determining whether the significant proportion of "Dancing with the Stars" watchers in the US" might confirm that the show was confusing and what that might mean. My first step was clean the data by filtering out between which show the response was for; then taking the cleaned data, I applied a bootstrap of 10,000 trials whilst deploying a confidence interval of 95% to observe if there is a significant difference of how confusing to the audience the show actually was. The bootstrap showed that after 10,000 trials the confidence interval ranged from 0.039 to 0.116 which does not include the value 0. Since our interval does not have the value 0 we can say there is a statistically significant difference of of people who agree that the show was confusing, a claim that can say we are 95% that between 3% and 12% of viewers thought "Dancing with the Stars" was confusing.

## Question 4

```{r, echo=FALSE, message=FALSE}

eb<- read_csv("ebay.csv", show_col_types = FALSE)
eb_ratio <- mutate(eb, rev_ratio = rev_after / rev_before)

eb_boot <- do(10000)*diffmean(rev_ratio ~ adwords_pause, data = resample(eb_ratio))
confint(eb_boot, level = 0.95)
```

The question tasked today is determining whether for both treatment and control groups in an experiment around pausing or not pausing for paid advertising within Google AdWords that the revenue ratio is constant and dose not change. My first step was clean the data by filtering out which entries paused or didn't pause the advertising; then taking the cleaned data, I applied a bootstrap of 10,000 trials whilst deploying a confidence interval of 95% to observe if there is a significant difference of how confusing to the audience the show actually was. The bootstrap showed that after 10,000 trials the confidence interval ranged from -0.090 to -0.013 which does not include the value 0 and a difference in mean revenue ratio of -0.052. Since our interval does not have the value 0 we can say there is a statistically significant difference of having paid search advertising. The data supports the idea that stopping paid search advertising decreases the revenue ratio, so instead eBay’s Google AdWords campaigns actively create additional profits. In general, for Ebay they should continue with the paid advertising since it appears to boost revenue by a little over 5%.



